Some examples in extended logic counting classes

### 2016-12-31 LICS/syntactic.tex
sin comentarios

The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a difficult problem. More specifically, it was shown in  \cite{Valiant79}  that this problem is $\shp$-complete. As a consequence of this result the problem of computing the number of perfect matchings of a bipartite graph was also shown to be $\shp$-complete. Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,BW05,DS12, PS13,PS14}. Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard. A first prominent example of such problems is counting the number of perfect matching of a bipartite graph, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time. Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula \cite{V79b}, satisfying assignments of a DNF propositional formula \cite{DHK05}, simple paths from a source node to a target node in a directed graph \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.

@@

The importance of the class counting problems with easy decision counterparts has motivated the search of robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}. But the key question here is what should be considered a {\em robust} class. A first condition to be met is that the class should have a syntactic definition that allows for the existence of natural complete problems. Special attention has to be paid here to the notion of reduction used for completeness. Notice that under the notion of Cook reduction, originally used in \cite{Valiant79}, the problems $\cdnf$ and $\csat$ are $\shp$-complete. However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$. A second condition to be met has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}. As in the cases of $\ptime$ and $\np$ that are closed under intersection and union, we expect our class to be closed under multiplication and sum; in fact, if $f$ and $g$ are counting functions with easy decision versions, then we clearly expect $f + g$ and $f \cdot g$ to have easy decision counterparts. For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula. The decision counterpart of such function is the language of all proposicional formulas $\varphi$ such that $\textit{sat\_one}(\varphi) > 0$. Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart is trivial. But should $\textit{sat\_one}$ be part of a robust class of counting function with easy decision versions? The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for $x$. As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula. Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them. In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g - 1$ should also be in $\CC$. Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} - 1$ counts the number of satisfying assignments of a propositional formula.

>>>>

The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a difficult problem. More specifically, it was shown in  \cite{Valiant79}  that this problem is $\shp$-complete. As a consequence of this result the $\shp$-completeness of the problem of computing the number of perfect matchings of a bipartite graph was established.
Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,BW05,DS12, PS13,PS14}. Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard. Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$. A first prominent example of such functions is counting the number of perfect matching of a bipartite graph, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time. Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula \cite{V79b}, satisfying assignments of a DNF propositional formula \cite{DHK05}, simple paths from a source node to a target node in a directed graph \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.

@@

The importance of the class counting problems with easy decision counterparts has motivated the search of robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}. But the key question here is what should be considered a {\em robust} class. A first desirable condition is that the class has a syntactic definition that allows for the existence of natural complete problems \cite{P94}. Special attention has to be paid here to the notion of reduction used for completeness. Notice that under the notion of Cook reduction, originally used in \cite{Valiant79}, the problems $\cdnf$ and $\csat$ are $\shp$-complete. However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$. Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfy that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$. 
A second desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}. As in the cases of $\ptime$ and $\np$ that are closed under intersection and union, we expect our class to be closed under multiplication and sum; in fact, if $f$ and $g$ are counting functions with easy decision versions, then we clearly expect $f + g$ and $f \cdot g$ to have easy decision counterparts. For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula. Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}} = \{ \varphi \mid \textit{sat\_one}(\varphi) > 0 \}$ is trivial. But should $\textit{sat\_one}$ be part of a robust class of counting function with easy decision versions? The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for $x$. As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula. Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them. In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g - 1$ should also be in $\CC$. Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} - 1$ counts the number of satisfying assignments of a propositional formula.
